---
title: "Lecture 2"
author: "BSS Stat 20"
date: '2022-06-22'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Constructing a Claim

The first thing we should do is load in the packages and data we will be using.

```{r, echo = TRUE, message = FALSE}
# load libraries

library(stat20data) #has the data we need
library(tidyverse)  # has the functions we will use

# load in the data

data(ppk)
```

Let's take a look at the data using the viewer in the top-right box of RStudio.

You can also run the following line.

```{r, echo = TRUE, eval = FALSE}
View(ppk)

```

We do have *levels* of support and opposition to the claim, but now we need to combine them into one support column and one opposition column. We do this using the `mutate()` function in the `dplyr` package, which is a part of the `tidyverse` library we loaded.

```{r}
#You will need to know of the mutate function this week, but not the entirety of the syntax. This will be for next week.

ppk <- mutate(ppk, support_before = Q18_words %
              in%
                    c("Somewhat support", "strongly support",
                      "Very strongly support"))

ppk <- mutate(ppk, opp_before = Q18_words %in%
                    c("Somewhat oppose", "strongly oppose",
                      "Very strongly oppose"))
```

Now here is a neat little trick to get the percentages that will be clearer to you next week.

```{r}
summarise(ppk, mean(support_before))

summarise(ppk, mean(opp_before))

```
