---
title: 'Lab 1: Introduction'
author: "BSS Stat 20"
date: 'June 22nd, 2022'
output:
  xaringan::moon_reader:
    css: stat20.css
    lib_dir: assets/libs
    nature:
      highlightStyle: atelier-forest-light
      highlightLines: true
      highlightSpans: true
      ratio: 16:9
      countIncrementalSlides: false
      slideNumberFormat: "%current%"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE,
                      warning = FALSE,
                      echo = FALSE,
                      fig.align = "center",
                      fig.retina = 3)

library(tidyverse)
library(xaringanthemer)
```

class: center, middle

# Welcome to Lab!

---

## Stat 20 Lab
.pull-left[
**Who**: 25-30 students + GSI

**What**: applying the concepts from lecture by working with real data  

**When**: Monday, Wednesday, Thursday afternoons; 50 minutes  

**Where**: Evans Hall 342  
]
--

.pull-right[

**Why**: Doing statistics on a data set requires a synthesis of conceptual frameworks, ideas and notation from mathematics, computation tools, and an understanding of the context of the data itself. It's a complex task and it requires practice.  

**How**: 
-   Groupwork and boardwork
-   Coding with R

]

---

## Icebreaker

1.  Get into groups of 4. Introduce yourselves with names, preferred pronouns, and hometowns.

2. Come up with *two truths and a lie*! Everyone guess each others' lies.

3.  Post your lie to an Ed forum we will create. It will be titled "[Your Section Number] Ed Practice"

---
## Lab 1 Goals

In completing this lab:

-   **Conceptually**, you will learn to:

    -   Identify the unit of observation in a data set
    -   Identify variables in a data set using the Taxonomy of Data

--

-  **Technincally**, you will be learn to:

    -   Use RMarkdown to compose documents with both images and R code
    -   Create vectors in R using `c()` and data frames using `tibble()`
    -   Explore a data set using some `dplyr` functions

---
class: center, middle
# Question One

Go ahead and visit [stat20.datahub.berkeley.edu](stat20.datahub.berkeley.edu)!
---
## The Big Three of Stat 20
1. R
  -   A programming language designed with statisticians in mind
  
--

2. RStudio
  -   A nice "wrapper" software to write R code in
  -   You can open an RStudio session by visiting [stat20.datahub.berkeley.edu](stat20.datahub.berkeley.edu)!
  
--

3. RMarkdown (.Rmd files)
  - A useful document format which combines standard word processing with R code chunks
  - To submit your labs, you will compose a .Rmd file, *knit* it to a PDF, and then submit
  
--
  
---
## The Components of RStudio

RStudio separates itself into four blocks, or *panes*. 

-  You will work most of the time in the **editor**, which is the top-left pane and the one which is most prominent on the screen. This is where you can compose .R and .Rmd files.

--

-  If you want to run R commands without a file, use the **console**, the bottom-left pane. Begin typing at the `>` and press enter to run your command.

---
## The Components of RStudio

-  When you make **objects** or load in **data frames**, they will appear in the top-right pane called the **environment.**

--

- Finally, the bottom-right pane is called the **file directory**. This is where all the files in your DataHub appear-- both the ones you make in RStudio and the ones you upload (a sketch of a data set, for example). 

---
## Running your first line of code
-  Open a new R script. This is a file that contains R code. 
  -   File -> New File -> R Script

--

-  Write a simple command. R is at its core a calculator. How about 5 + 3?

--

-  Select the entire line of code with your cursor. To run the code, either:
  -   Click the *"Run" Button* in the top right corner of the edtior
  -   *Control + Enter* on a Windows computer 
  -   *Command + Enter* on a Mac

--

-  Once you do this, you should see the result (8) pop up in the console!

---
## What can R do?

- We have seen that R can add with `+`. R can also subtract with `-`, multiply with `*`, and divide with `/`.

- R can also work with words, or *character strings*.

--

```{r, echo = TRUE}
"Jeremy"
```

---
## Exercise

- R follows the order of operations! Get back into your same groups of 3-5. You will be assigned one of these mathematical expressions. Come up with the R code to write it. You are free to consult any sources you need. 

1.  the quantity 5 multiplied by 2, raised to the power of 3, plus 1
  - The answer should be 1001   
  
2.  5 multipiled by the quantity 2 raised to the power 3, plus 1
  - The answer should be 41

- Then go ahead and write your group's answer somewhere on the board! 

---
## What can R do?

- What if we want to store a number or a character string? We use `<-`.

--

```{r, echo = TRUE}
favorite_number <- 20
favorite_number
```

--

```{r, echo = TRUE}
favorite_word <- "statistics"
favorite_word
```

--

- We call `favorite_number` and `favorite_word` **objects** that store a *value*.
- We call `<-` the **assignment operator**.
  - We would say that `20` was **assigned** to the **object** `favorite_number`.

---
## What can R do?

- What if I want to store multiple numbers or words at a time? We use `c()`.

- `c` is for["concatenate"](https://www.oxfordreference.com/view/10.1093/oi/authority.20110803095630306), which means to link things together end to end in a chain!

```{r, echo = TRUE}
favorite_numbers <- c(20, 21, 22, 23)
favorite_numbers
```


```{r, echo = TRUE}
course_staff <- c("Jeremy", "Wayne", "Qingyi", "Dev")
course_staff
```

---
## What can R do?

-   **Vectors** are a fundamental building block of data in R and here are some things we can do with them.

-   For numeric vectors, we can use built-in R functions to calculate statistics.

--

```{r, echo = TRUE}
jeremys_graduation_years <- c(2013, 2017, 2021, 2022)
mean(jeremys_graduation_years) #returns the average
sum(jeremys_graduation_years)  #returns the total
```


---
## What can R do?

- R can also store data sets. We call data sets in R **data frames**.

--

- We can either load existing data sets or we can make them from scratch.

---
## What can R do?

- Loading existing data sets (such as `iris`, a data set included in R)

--

```{r, echo = TRUE, message = FALSE}
#head prints the first six rows of the data set by default
head(iris)
```

---
## What can R do?

- Making data frames from scratch with `tibble()`!

--

```{r, echo = TRUE, message = FALSE}
#what is this library business?
library(tidyverse) 

course_staff <- c("Jeremy", "Wayne", "Qingyi", "Dev")
favorite_numbers <- c(20, 21, 22, 23)

first_data_frame <- tibble(course_staff, favorite_numbers)

#Renaming the columns of our data frame with a new name vector!
names(first_data_frame) <- c("Staff", "Numbers")

first_data_frame
```

---
## What can R do (with help)?

-   Lastly, what did `library(tidyverse)` do in the last slide?

--

-   `tibble()` is a function in a **package** called `dplyr`, which itself is a part of the `tidyverse` **library**

--

-   **Packages** (and the **libraries** they are sometimes inside of), offer extra functionality that doesn't come with the base installation of R

---
## What can R do (with help)?

-   Because the packages don't come automatically, we have to *load them in* (just once) using `library(PACKAGE_NAME_HERE)`

--

-   We will be using a few different libraries/packages this semester!

    -   The `tidyverse` library
    
--

        -   The `dplyr` package for data analysis and wrangling
--
        
        -   The `ggplot2` package for data visualization
--

    -   The `stat20data` package includes data sets we will use
--

    -   The `infer` package for inferential statistics
  
---
## Question 1

With all of this in hand, you should be able to finish Question 1 now and be prepared for the remainder of the lab!



